A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$ methods

### 2024-10-28 paper_v01/article.tex
update paper

%% - \Abstract{} almost as usual
\Abstract{
  This short article illustrates how to write a manuscript for the
  \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
  Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
  it is recommended to keep the {\LaTeX} code as simple as possible,
  i.e., avoid inclusion of packages/commands that are not necessary.
  For outlining the typical structure of a JSS article some brief text snippets
  are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
  discussing count data regression in \proglang{R}. Editorial comments and
  instructions are marked by vertical bars.
}

@@

The ${DMC}_{x}^{2}$ ranges from zero $(0)$, where the time-series do not correlate, to one $(1)$ meaning that the time-series have either a perfect correlation or a perfect anti-correlation. In this equation, $\rho^{-1}(n)$ represent the inverse matrix of all possible combinations of $\rho_{DCCA}$ between the independent variables, in other words:

@@

Modeling count variables is a common task in economics and the social sciences.
The classical Poisson regression model for count data is often of limited use in
these disciplines because empirical count data sets typically exhibit
overdispersion and/or an excess number of zeros. The former issue can be
addressed by extending  the plain Poisson regression model in various
directions: e.g., using sandwich covariances or estimating an additional
dispersion parameter (in a so-called quasi-Poisson model). Another more formal
way is to use a negative binomial (NB) regression. All of these models belong to
the family of generalized linear models (GLMs). However, although these models
typically can capture overdispersion rather well, they are in many applications
not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
increased interest in zero-augmented models that address this issue by a second
model component capturing zero counts. An overview of count data models in
econometrics, including  hurdle and zero-inflated models, is provided in
\cite{Cameron+Trivedi:2013}.

In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
diagnostics and inference. The manuscript that this document is based on
\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
package \citep{Jackman:2015}. Of course, much more software could be discussed
here, including (but not limited to) generalized additive models for count data
as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006},
\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.

@@

The basic Poisson regression model for count data is a special case of the GLM
framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
log link and a linear predictor
%
\begin{equation} \label{eq:mean}
  \log(\mu_i) \quad = \quad x_i^\top \beta,
\end{equation}
%
where the regression coefficients $\beta$ are estimated by maximum likelihood
(ML) using the iterative weighted least squares (IWLS) algorithm.

\begin{leftbar}
  Note that around the \verb|{equation}| above there should be no spaces (avoided
  in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
  not a new paragraph started.
\end{leftbar}

\proglang{R} provides a very flexible implementation of the general GLM
framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
\pkg{stats} package. Its most important arguments are
\begin{Code}
  glm(formula, data, subset, na.action, weights, offset,
  family = gaussian, start = NULL, control = glm.control(...),
  model = TRUE, y = TRUE, x = FALSE, ...)
\end{Code}
where \code{formula} plus \code{data} is the now standard way of specifying
regression relationships in \proglang{R}/\proglang{S} introduced in
\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
standard  for setting up formula-based regression models in
\proglang{R}/\proglang{S}. The arguments in the second line control aspects
specific to GLMs while the arguments in the last line specify which components
are returned in the fitted model object (of class \class{glm} which inherits
from \class{lm}). For further arguments to \fct{glm} (including alternative
specifications of starting values) see \code{?glm}. For estimating a Poisson
model \code{family = poisson} has to be specified.

\begin{leftbar}
  As the synopsis above is a code listing that is not meant to be executed,
  one can use either the dedicated \verb|{Code}| environment or a simple
  \verb|{verbatim}| environment for this. Again, spaces before and after should be
  avoided.

  Finally, there might be a reference to a \verb|{table}| such as
  Table~\ref{tab:overview}. Usually, these are placed at the top of the page
  (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
  column headers and captions in sentence style, and if possible avoiding vertical
  lines.
\end{leftbar}

\begin{table}[t!]
  \centering
  \begin{tabular}{lllp{7.4cm}}
    \hline
    Type           & Distribution & Method   & Description                        \\ \hline
    GLM            & Poisson      & ML       & Poisson regression: classical GLM,
    estimated by maximum likelihood (ML)                                          \\
                   &              & Quasi    & ``Quasi-Poisson regression'':
    same mean function, estimated by
    quasi-ML (QML) or equivalently
    generalized estimating equations (GEE),
    inference adjustment via estimated
    dispersion parameter                                                          \\
                   &              & Adjusted & ``Adjusted Poisson regression'':
    same mean function, estimated by
    QML/GEE, inference adjustment via
    sandwich covariances                                                          \\
                   & NB           & ML       & NB regression: extended GLM,
    estimated by ML including additional
    shape parameter                                                               \\ \hline
    Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
    hurdle Poisson                                                                \\
                   & NB           & ML       & Zero-inflated NB (ZINB),
    hurdle NB                                                                     \\ \hline
  \end{tabular}
  \caption{\label{tab:overview} Overview of various count regression models. The
    table is usually placed at the top of the page (\texttt{[t!]}), centered
    (\texttt{centering}), has a caption below the table, column headers and captions
    are in sentence style, and if possible vertical lines should be avoided.}
\end{table}

@@

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> data("quine", package = "MASS")
  \end{CodeInput}
\end{CodeChunk}
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

\begin{leftbar}
  For code input and output, the style files provide dedicated environments.
  Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
  or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
  produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
  hook. Please make sure that all code is properly spaced, e.g., using
  \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
  use ``the usual'' command prompt in the respective software system. For
  \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
  the continuation prompt. Generally, comments within the code chunks should be
  avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
  before and after code input/output should be avoided (see above).
\end{leftbar}

\begin{figure}[t!]
  \centering
  \includegraphics{article-visualization}
  \caption{\label{fig:quine} Frequency distribution for number of days absent
    from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
    +    family = poisson)
  \end{CodeInput}
\end{CodeChunk}
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> library("MASS")
    R> m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
  \end{CodeInput}
\end{CodeChunk}
%
In a comparison with the BIC the latter model is clearly preferred.
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> BIC(m_pois, m_nbin)
  \end{CodeInput}
  \begin{CodeOutput}
    df      BIC
    m_pois 18 2046.851
    m_nbin 19 1157.235
  \end{CodeOutput}
\end{CodeChunk}
%
Hence, the full summary of that model is shown below.
%
\begin{CodeChunk}
  \begin{CodeInput}
    R> summary(m_nbin)
  \end{CodeInput}
  \begin{CodeOutput}
    Call:
    glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
    init.theta = 1.60364105, link = log)

    Deviance Residuals:
    Min       1Q   Median       3Q      Max
    -3.0857  -0.8306  -0.2620   0.4282   2.0898

    Coefficients: (1 not defined because of singularities)
    Estimate Std. Error z value Pr(>|z|)
    (Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
    EthN        -0.24591    0.39135  -0.628  0.52977
    SexM        -0.77181    0.38021  -2.030  0.04236 *
    AgeF1       -0.02546    0.41615  -0.061  0.95121
    AgeF2       -0.54884    0.54393  -1.009  0.31296
    AgeF3       -0.25735    0.40558  -0.635  0.52574
    LrnSL        0.38919    0.48421   0.804  0.42153
    EthN:SexM    0.36240    0.29430   1.231  0.21818
    EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876
    EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 **
    EthN:AgeF3   0.04721    0.44883   0.105  0.91622
    EthN:LrnSL   0.06847    0.34040   0.201  0.84059
    SexM:AgeF1   0.02257    0.47360   0.048  0.96198
    SexM:AgeF2   1.55330    0.51325   3.026  0.00247 **
    SexM:AgeF3   1.25227    0.45539   2.750  0.00596 **
    SexM:LrnSL   0.07187    0.40805   0.176  0.86019
    AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870
    AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363
    AgeF3:LrnSL       NA         NA      NA       NA
    ---
    Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

    (Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)

    Null deviance: 235.23  on 145  degrees of freedom
    Residual deviance: 167.53  on 128  degrees of freedom
    AIC: 1100.5

    Number of Fisher Scoring iterations: 1

    Theta:  1.604
    Std. Err.:  0.214

    2 x log-likelihood:  -1062.546
  \end{CodeOutput}
\end{CodeChunk}

@@

\section*{Computational details}

\begin{leftbar}
  If necessary or useful, information about certain computational details
  such as version numbers, operating systems, or compilers could be included
  in an unnumbered section. Also, auxiliary packages (say, for visualizations,
  maps, tables, \dots) that are not cited in the main text can be credited here.
\end{leftbar}

The results in this paper were obtained using
\proglang{R}~3.4.1 with the
\pkg{MASS}~7.3.47 package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

\begin{leftbar}
  All acknowledgments (note the AE spelling) should be collected in this
  unnumbered section before the references. It may contain the usual information
  about funding and feedback from colleagues/reviewers/etc. Furthermore,
  information such as relative contributions of the authors may be added here
  (if any).
\end{leftbar}

@@

\newpage

\begin{appendix}

  \section{More technical details} \label{app:technical}

  \begin{leftbar}
    Appendices can be included after the bibliography (with a page break). Each
    section within the appendix should have a proper section title (rather than
    just \emph{Appendix}).

    For more technical style details, please check out JSS's style FAQ at
    \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
    which includes the following topics:
    \begin{itemize}
      \item Title vs.\ sentence case.
      \item Graphics formatting.
      \item Naming conventions.
      \item Turning JSS manuscripts into \proglang{R} package vignettes.
      \item Trouble shooting.
      \item Many other potentially helpful details\dots
    \end{itemize}
  \end{leftbar}

  \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

  \begin{leftbar}
    References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
    references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
    \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
    formats of author-year citations and allow to include additional details (e.g.,
    pages, chapters, \dots) in brackets. In case you are not familiar with these
    commands see the JSS style FAQ for details.

    Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
    when acquiring the entries automatically from mixed online sources. However,
    it is important that informations are complete and presented in a consistent
    style to avoid confusions. JSS requires the following format.
    \begin{itemize}
      \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
            be used in the references.
      \item Titles should be in title case.
      \item Journal titles should not be abbreviated and in title case.
      \item DOIs should be included where available.
      \item Software should be properly cited as well. For \proglang{R} packages
            \code{citation("pkgname")} typically provides a good starting point.
    \end{itemize}
  \end{leftbar}

\end{appendix}

>>>>

%% - \Abstract{} almost as usual
\Abstract{
  This paper presents tha \pkg{Zebende}, a \proglang{Python} package writen in \proglang{Python} and \proglang{Zig}, that calculates the $DFA$, $DCCA$ $\rho_{DCCA}$ and the $DMC_x^2$. The package presents an optimized algorithm that significantly improves the calculations speed. A comparison with ohter packages that calculates the .The package is also the firts to implement the $DMC_x^2$ corefficient for any number of series.
}

@@

@@

@@

@@

@@

% \section*{Computational details}

% \begin{leftbar}
%   If necessary or useful, information about certain computational details
%   such as version numbers, operating systems, or compilers could be included
%   in an unnumbered section. Also, auxiliary packages (say, for visualizations,
%   maps, tables, \dots) that are not cited in the main text can be credited here.
% \end{leftbar}

% The results in this paper were obtained using
% \proglang{R}~3.4.1 with the
% \pkg{MASS}~7.3.47 package. \proglang{R} itself
% and all packages used are available from the Comprehensive
% \proglang{R} Archive Network (CRAN) at
% \url{https://CRAN.R-project.org/}.

% \section*{Acknowledgments}

% \begin{leftbar}
%   All acknowledgments (note the AE spelling) should be collected in this
%   unnumbered section before the references. It may contain the usual information
%   about funding and feedback from colleagues/reviewers/etc. Furthermore,
%   information such as relative contributions of the authors may be added here
%   (if any).
% \end{leftbar}

@@

% \newpage

% \begin{appendix}

%   \section{More technical details} \label{app:technical}

%   \begin{leftbar}
%     Appendices can be included after the bibliography (with a page break). Each
%     section within the appendix should have a proper section title (rather than
%     just \emph{Appendix}).

%     For more technical style details, please check out JSS's style FAQ at
%     \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
%     which includes the following topics:
%     \begin{itemize}
%       \item Title vs.\ sentence case.
%       \item Graphics formatting.
%       \item Naming conventions.
%       \item Turning JSS manuscripts into \proglang{R} package vignettes.
%       \item Trouble shooting.
%       \item Many other potentially helpful details\dots
%     \end{itemize}
%   \end{leftbar}

%   \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

%   \begin{leftbar}
%     References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
%     references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
%     \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
%     formats of author-year citations and allow to include additional details (e.g.,
%     pages, chapters, \dots) in brackets. In case you are not familiar with these
%     commands see the JSS style FAQ for details.

%     Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
%     when acquiring the entries automatically from mixed online sources. However,
%     it is important that informations are complete and presented in a consistent
%     style to avoid confusions. JSS requires the following format.
%     \begin{itemize}
%       \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
%             be used in the references.
%       \item Titles should be in title case.
%       \item Journal titles should not be abbreviated and in title case.
%       \item DOIs should be included where available.
%       \item Software should be properly cited as well. For \proglang{R} packages
%             \code{citation("pkgname")} typically provides a good starting point.
%     \end{itemize}
%   \end{leftbar}

% \end{appendix}
